Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-05-15 draft.tex
added/modified few things.

There are two processes to consider in order to estimate the rate of energy loss via $\Aprime$ bremsstrahlung. The first is  
proton-proton (pp) scattering with the bremsstrahlung of the dark photon off the proton; $p+p \rightarrow p+p+\Aprime$. 
The second is proton-neutron (pn) scattering with bremsstrahlung off of either the proton or the charged pion; 
$p+n \rightarrow p+n+\Aprime$. We estimate the rates for these processes using the one-pion exchange (OPE) 
approximation for nucleon interactions. In the pp case, there are eight tree-level diagrams, 
with the emission of the $\Aprime$ from each of the external legs. One of these diagrams is shown 
in Fig.~\ref{fig:ppdiagram}; the remaining seven diagrams come from placing the radiated $\Aprime$ 
on each of the other three protons and then, for each of these, interchanging the outgoing momenta. 
For the pn case, there are five diagrams, four of which are analogous to the pp diagram shown in 
Fig.~\ref{fig:ppdiagram}. The fifth diagram, shown in Fig.~\ref{fig:npdiagram}, corresponds to 
emission of the $\Aprime$ from the exchanged, charged pion. \akl{Cameron, can you make the $A' \to \mathrm{A'}$ in figures 1 and 2?}

@@

The inverse mean free path can be obtained by modifying $ Q_i $, 
the expression for the energy loss rate, as follows: removing the factor of $ E_\mathrm{A'} $ 
and the phase space integral over $ \qa $, and adding a factor of $ e^{E_\mathrm{A'}/T} $ for detailed balance. 
\arz{I'm not sure that I follow this somewhat cavalier statement about detailed balance. Can someone 
please explain this to me when we meet. In my earlier rounds of thinking about this, I took that factor 
at face value as given, assuming I'd eventually be able to figure out where it comes from.} 
This gives the inverse mean free path as a function of mass and coupling. 
Again, we perform the required integration numerically using the {\tt CUBA }package. 
This then allows the calculation of $\kappa_{\mathrm{pp}}$ and $\kappa_{\mathrm{pn}}$, 
which are the opacities due to inverse bremsstrahlung for proton-proton and proton-neutron 
processes respectively. These dominate the opacity of the star to $\mathrm{A'}$ propagation. 
The inverse opacities for the $\mathrm{pn}$ and $\mathrm{pp}$ processes add, 
giving the total opacity $ \kappa^{-1} = \kappa_{\mathrm{pp}}^{-1} + \kappa_{\mathrm{pn}}^{-1} $

Having obtained an expression for $ \kappa $, we can now find the optical depth as follows. 
Define a typical optical depth as $\tau_\mathrm{R} = \kappa_\mathrm{R} \rho_\mathrm{R} R $. We then have 
\beq 
\kappa \rho R = \tau_\mathrm{R} \left( \frac{\rho}{\rho_\mathrm{R}} \right)^2 \left( \frac{T_\mathrm{R}}{T} \right)^{3/2}.
\eeq
%
This can be combined with the previous expressions for the density and temperature as a function of position 
and plugged in to the integral expression for the optical depth to obtain 
\bea 
\tau &=& \int_{r_x}^{\infty}\, \tau_\mathrm{R} \left( \frac{R}{r} \right)^{3n/2} \dd r \\
 &=& \frac{\tau_\mathrm{r}}{\frac{3n}{2}-1} \left( \frac{T_\mathrm{A'}}{T_\mathrm{R}} \right)^{(9/2-3/n)}. 
 \eea
 The bound on the coupling is obtained by requiring  $ \tau(\varepsilon,m_\mathrm{A'}) \le 2/3 $. Note that strictly speaking we should have determined $ r_{\mathrm{dp}}$ rather than assuming a value. However, it is possible to verify the self-consistency of our assumption using this model for the supernova atmosphere. In particular, this model implies that an optical depth of $\tau=2/3$ is reached at a radial position of $ r_{\mathrm{dp}}=XXXX$,  validating the assumption made at the outset. \arz{Put in the missing number here.} As we stated in the previous subsection, the approximate treatment of the upper limit of the exclusion region is justified by the fact that values of $\varepsilon$ close to the upper limit of the exclusion region are ruled out by independent, terrestrial experiments.

@@

\begin{figure*}[th]
\includegraphics[width=14cm]{compiledbounds_edited.pdf}
\caption{
Constraints on dark photon models. Our primary result is the excluded region 
in the $\varepsilon$-$m_{\mathrm{A'}}$ plane shaded in blue. The red curve is our attempts to reproduce exactly Ref.~\cite{dent_etal12}, while the black curve is Ref.~\cite{dent_etal12} but with the correct luminosity constraint.  The purple curve is using the SRA fixing the cross section using nuclear scattering data, while the green curve uses the soft radiation approximation of the OPE result.  \arz{The labels in the figure legend appear to be incorrect now.} \akl{In the legend, the red and black lines need to be switched, and the "Full excluded region" needs to be switched with the "SRA using OPE model".}
}
\label{fig:constraint}
\end{figure*}

>>>>

There are two processes to consider in order to estimate the rate of energy loss via $\Aprime$ bremsstrahlung. The first is  
proton-proton (pp) scattering with the bremsstrahlung of the dark photon off the proton; $p+p \rightarrow p+p+\Aprime$. 
The second is proton-neutron (pn) scattering with bremsstrahlung off of either the proton or the charged pion; 
$p+n \rightarrow p+n+\Aprime$. We estimate the rates for these processes using the one-pion exchange (OPE) 
approximation for nucleon interactions. In the pp case, there are eight tree-level diagrams, 
with the emission of the $\Aprime$ from each of the external legs. One of these diagrams is shown 
in Fig.~\ref{fig:ppdiagram}; the remaining seven diagrams come from placing the radiated $\Aprime$ 
on each of the other three protons and then, for each of these, interchanging the outgoing momenta. 
For the pn case, there are five diagrams, four of which are analogous to the pp diagram shown in 
Fig.~\ref{fig:ppdiagram}. The fifth diagram, shown in Fig.~\ref{fig:npdiagram}, corresponds to 
emission of the $\Aprime$ from the exchanged, charged pion. \akl{Andrew, can you make the $A' \to \mathrm{A'}$ in figures 1 and 2?}

@@

The inverse mean free path can be obtained by modifying $ Q_i $, 
the expression for the energy loss rate, as follows: removing the factor of $ E_\mathrm{A'} $ 
and the phase space integral over $ \qa $, and adding a factor of $ e^{E_\mathrm{A'}/T} $ for detailed balance. This factor comes from turning the $\mathrm{A'}$ from an outgoing to an incoming state in the calculation.  See, e.g., Eq.~(4.43) in \cite{raffelt96_book}.
\arz{I'm not sure that I follow this somewhat cavalier statement about detailed balance. Can someone 
please explain this to me when we meet. In my earlier rounds of thinking about this, I took that factor 
at face value as given, assuming I'd eventually be able to figure out where it comes from.} \akl{Is that good enough?}
This gives the inverse mean free path as a function of mass and coupling. 
Again, we perform the required integration numerically using the {\tt CUBA }package. 
This then allows the calculation of $\kappa_{\mathrm{pp}}$ and $\kappa_{\mathrm{pn}}$, 
which are the opacities due to inverse bremsstrahlung for proton-proton and proton-neutron 
processes respectively. These dominate the opacity of the star to $\mathrm{A'}$ propagation. 
The inverse opacities for the $\mathrm{pn}$ and $\mathrm{pp}$ processes add, 
giving the total opacity $ \kappa^{-1} = \kappa_{\mathrm{pp}}^{-1} + \kappa_{\mathrm{pn}}^{-1} $

Having obtained an expression for $ \kappa $, we can now find the optical depth as follows. 
Define a typical optical depth as $\tau_\mathrm{R} = \kappa_\mathrm{R} \rho_\mathrm{R} R $. We then have 
\beq 
\kappa \rho R = \tau_\mathrm{R} \left( \frac{\rho}{\rho_\mathrm{R}} \right)^2 \left( \frac{T_\mathrm{R}}{T} \right)^{3/2}.
\eeq
%
This can be combined with the previous expressions for the density and temperature as a function of position 
and plugged in to the integral expression for the optical depth to obtain 
\bea 
\tau &=& \int_{r_x}^{\infty}\, \tau_\mathrm{R} \left( \frac{R}{r} \right)^{3n/2} \dd r \\
 &=& \frac{\tau_\mathrm{r}}{\frac{3n}{2}-1} \left( \frac{T_\mathrm{A'}}{T_\mathrm{R}} \right)^{(9/2-3/n)}. 
 \eea
 The bound on the coupling is obtained by requiring  $ \tau(\varepsilon,m_\mathrm{A'}) \le 2/3 $. Note that strictly speaking we should have determined $ r_{\mathrm{dp}}$ rather than assuming a value. However, it is possible to verify the self-consistency of our assumption using this model for the supernova atmosphere. In particular, this model implies that an optical depth of $\tau=2/3$ is reached at a radial position of $ r_{\mathrm{dp}}= 11$ km,  validating the assumption made at the outset. As we stated in the previous subsection, the approximate treatment of the upper limit of the exclusion region is justified by the fact that values of $\varepsilon$ close to the upper limit of the exclusion region are ruled out by independent, terrestrial experiments.

@@

\begin{figure*}[th]
\includegraphics[width=14cm]{compiledbounds_edited.pdf}
\caption{
Constraints on dark photon models. Our primary result is the excluded region 
in the $\varepsilon$-$m_{\mathrm{A'}}$ plane shaded in blue. The red curve is our attempts to reproduce exactly Ref.~\cite{dent_etal12}, while the black curve is Ref.~\cite{dent_etal12} but with the correct luminosity constraint.  The purple curve is using the SRA fixing the cross section using nuclear scattering data, while the green curve uses the soft radiation approximation of the OPE result.  \arz{The labels in the figure legend appear to be incorrect now.} \akl{In the legend, the red and black lines need to be switched, and the "Full excluded region" needs to be switched with the "SRA using OPE model". Also, can you put a cdot or space between erg and s in the vertical label.}
}
\label{fig:constraint}
\end{figure*}
